
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style

-------

Questions
-------

1. Create the data frame

```{r}
df = data.frame(Rank = c(0, 1, 2, 3),
                Age = c(28, 34, 29, 42),
                Name = c('Tom', 'Jack', 'Steve', 'Ricky'))
df
```


2. Use read.csv to import the Covid19 Vaccination data
```{r}
setwd("C:/Users/student/Documents/GitHub/MATH421")
df = read.csv('us-states.csv')
```


Show the names of the variables in the data

```{r}
names(df)
```


3. How many columns and rows the data have? 

```{r}
str(df)

# There are five columns and 50,878 rows. 
```


4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?

```{r}
sum(is.na(df))
colSums(is.na(df))
colMeans(is.na(df))

# there are no missing values
```


5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 
```{r}
class(df$date)
df$date = as.Date(df$date)
str(df)
```


6. What is the class of the date column. Change the date columns to date type using the as.Date function. Show the new class of the date column.
```{r}
class(df$Date)

names(df) [1] = 'Date'
names(df) [2] = 'State'
names(df) [3] = 'Fips'
names(df) [4] = 'Cases'
names(df) [5] = 'Deaths'
names(df)

class(df$Date)
```


7. Find the average number of cases per day. Find the maximum cases a day.  
```{r}
mean(df$Cases, na.rm=TRUE)
max(df$Cases, na.rm=TRUE)
```


8. How many states are there in the data?  
```{r}
length(table(df$State))
```


9. Create a new variable `weekdays` to store the weekday for each rows. 
```{r}
df$weekdays <- weekdays(df$Date)
```


10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day
```{r}
df$Deaths2 = ifelse(df$Deaths==0, "No", "Yes")
```


Find the frequency and relative frequency of `no_death` and `has_death`. 

11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)
```{r}
summary(df$Deaths)
```


12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3
```{r}
library(dplyr)
df$Death3 = case_when(df$Deaths <4046 ~ "Low", 
                      df$Deaths <12674 ~ "Medium", 
                      TRUE ~ "High" )
```


13. Find the average cases in Rhode Island in 2021
```{r}
df1 = df[df$Date>'2021-01-01',] 
by (df$Cases, df$State, mean)
```


14. Find the median cases by weekdays in Rhode Island in 2021 
```{r}
df$weekdays <- weekdays(df$Date)
df1 = df[df$State=='Rhode Island',]
by(data = df1$Deaths, INDICES = df1$weekdays, FUN = median)
```


15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 
```{r}
df$month <- months(df$Date)

df$month <- months(df$Date)
df1 = df[df$State=='Rhode Island',]
by(data = df1$Deaths, INDICES = df1$month, FUN = median)
```


16.  Find your own dataset, import it and implement the following functions on the data

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when
```{r}
setwd("C:/Users/student/Documents/GitHub/MATH421")
read.csv("titanic (2).csv")
dft = read.csv("titanic (2).csv")

head(dft)
str(dft)
names(dft)
sum(is.na(dft))
colSums(is.na(dft))
colMeans(is.na(dft))
mean(dft$Survived, na.rm=TRUE)
max(dft$Survived, na.rm=TRUE)
cor(dft$Fare, dft$SibSp)
by(dft$Survived, dft$Cabin, mean)
dft$Survived1 <- ifelse(dft$Survived==0, 
                        'No_Survivors','Has_Survivors')
dft$Survived1 <- case_when(dft$Survived==0 ~ 'No_Survivors', 
                           TRUE ~ 'Has_Survivors')

summary(dft)

```


17.  In the dataset in #16, practice the follows.  You can reuse the code of 16. 

- Create a categorical variable from a continuous variable
- From one catego;rical variable, create a new categorical variable with fewer categories
```{r}
dft$Survived1 <- ifelse(dft$Survived==0, 
                        'No_Survivors','Has_Survivors')
dft$Embarked2 = ifelse(dft$Embarked==0, "Q", "S")

summary(dft)
```


If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)
